---
title: "igraph Package"
author: 'Asaf, Steve, Xin'
output: html_document
---
The R package 'igraph' allows for graphs to be generated using algorithms and 
improves the ability of R to handle extremely large graphs. Another powerful use of iGraph is the 
direct implementation of data frames as graphs. Primarily, igraph can be used for the analysis of network theory.



Dependencies: 
The igraph package requires that irlba and pkgconfig ##add more info



Data Set: The data used to demonstrate the functionality of igraph comes from networks of media outlets and their audience sizes.

Source: http://kateto.net/networks-r-igraph (from section 3.1)



Basic arguments of graph function:

--Types of basic plot functions:: There are several ways which a plot can be created in this package. One method uses 
 'graph_from_literal'. This function requires a formula which shows each vertex and its connection to other vertices in the plot.
 A useful feature of igraph is the ability to directly convert a data frame to a graph, using the 'graph_from_data_frame' function.
 Other functions are 'graph_from_edgelist', and 'graph_from_adjacency_matrix'.
 
--Vertices/Edges: Verticies are points within the graph, and edges are the lines which connect each vertex. The basic graph function
 allows the number of verticies to be specified. Alternatively, some igraph functions allow a dataframe or list of vertices and/or edges to  
 be passed into the function to create a graph.
 
--Directed/Undirected:: There are two main different types of iGraphs. The first is undirected and the other is directed; Undirected means
 that the edges don't specify a direction for the connection between verticies. Directed graphs specify a single path between verticies.
 By default, the 'directed' arguement is TRUE.
 
--Simplify: This argument takes TRUE or FALSE as its input (TRUE by default). If the value is TRUE, then the graph will not allow and edge 
 to connect both of its endpoints to a single vertex. A simplified graph also prevents multiple edges; this means that it will prevent two
 edges from having identical endpoints.


Basic iGraph functions:

-graph or make_graph

-vertex_attr

-degree

-add.edges

-add.vertices

-delete.edges

-delete.vertices


```{r, message=FALSE}
library(igraph)
group <- graph( edges=c('A','B', 'B','C', 'C','D', 'D','A'))
plot(group, main='Basic Graph')
```

Directed vs Undirected plots
```{r}
par(mfrow=c(1,2))

group1 <- graph( edges=c(1,2, 2,3, 3,4, 4,1), directed=F) 
plot(group1,main = 'Undirected')
group2 <- graph( edges=c(1,2, 2,3, 3,4, 4,1), directed=T)
plot(group2,main='Directed')
```



The parameter n can specify the number of vertices for a graph when the edges are numeric and will affect the plot
when n is at least equal to the number of edges specified
```{r}
par(mfrow=c(1,2))

group3 <- graph(edges=c(1,2, 2,3, 3,4, 4,1), n=6)
plot(group3)
group4 <- graph(edges=c(1,2, 2,3, 3,4, 4,1), n=3)
plot(group4) # produces same result as first graph
```

Demonstrations of Basic Functions:

-vertex attributes
```{r}

```


-add vertex/edge:
```{r}
par(mfrow=c(1,2))

group <- graph(edges=c(1,2, 2,3, 3,4, 4,1))
plot(group)
add.vertices(group, 2, color='red')
plot(group)

```


-delete vertex/edge:
```{r}
group <- graph(edges=c(1,2, 2,3, 3,4, 4,1), n=6)
plot(group)
```



Plot generated from a data frame:
```{r}
edges = data.frame(x = c('A','B','C','D'), y = c('B','C','D','A'))
graph1 <- graph_from_data_frame(edges)
plot(graph1,main='From Dataframe')
```


Plot generated from graph_from_literal:
```{r}
par(mfrow=c(1,2))

graph2 <- graph_from_literal( "+" -- "-", "*" -- "/", "%" -- "&" )
plot(graph2,main='From Literal')
graph2 <- graph_from_literal( "+" -+ "-", "*" -+ "/", "%" -+ "&" )
plot(graph2,main='From Literal - Directed')
```


Plot generated from edgelist:
```{r}
edgelist = matrix(c(1, 2, 3, 4, 5, 6), ncol = 2, byrow = TRUE)
edgelist
graph3 <- graph_from_edgelist(edgelist, directed = FALSE)
plot(graph3,main='From Edgelist')
```


Plot generated from adjacency matrix:
 Note: This matrix must be square. The dimensions of the adjacency matrix determine the number of vertices. The value of each coordinate 
 in the matrix determines 
```{r}
par(mfrow=c(1,2))
adjm <- matrix(rnorm(100,0,1), nrow=10, ncol=10)
graph4<- graph_from_adjacency_matrix(adjmatrix = adjm, mode="undirected")
plot(graph4,main='From Adjacency Matrix')
#origin matrix and adjacency matrix
adjm 
get.adjacency((graph4)) 
#simplified
graph4 <- simplify(graph4)
plot(graph4,main='From Adjacency Matrix - Simplified')
get.adjacency((graph4)) 

```
In the adjacency matrix, a 1 on the diagonal will generate a loop, while a number greater than 1 will show up as multiple edges between nodes.

Additional Example Plots:

```{r}
emptyGraph <- make_empty_graph(20)
plot(emptyGraph, vertex.size=20, main = 'Empty Graph')
```

```{r}
fullGraph <- make_full_graph(10)
plot(fullGraph, vertex.size=10, vertex.label=NA, main='Full Graph')
degree(fullGraph)
get.adjacency(fullGraph)
```

```{r}
starGraph <- make_star(20)
plot(starGraph, vertex.size=10, vertex.label=NA, main='Star Graph') 
degree(starGraph)
get.adjacency(starGraph)
```

```{r}
treeGraph <- make_tree(20, children = 3, mode = "undirected")
plot(treeGraph, vertex.size=10, vertex.label=NA, main='Tree Graph') 
get.adjacency(treeGraph)
```

```{r}
ringGraph <- make_ring(20)
plot(ringGraph, vertex.size=10, vertex.label=NA, main='Ring Graph')
degree(ringGraph)
get.adjacency(ringGraph)
```


#####



```{r, message = FALSE}
library(igraph)
links = read.csv('https://raw.githubusercontent.com/kateto/R-igraph-Network-Workshop-NetSciX/master/Dataset1-Media-Example-EDGES.csv')
nodes = read.csv('https://raw.githubusercontent.com/kateto/R-igraph-Network-Workshop-NetSciX/master/Dataset1-Media-Example-NODES.csv')
links2 = read.csv('https://raw.githubusercontent.com/kateto/R-igraph-Network-Workshop-NetSciX/master/Dataset2-Media-User-Example-EDGES.csv')
nodes2 = read.csv('https://raw.githubusercontent.com/kateto/R-igraph-Network-Workshop-NetSciX/master/Dataset2-Media-User-Example-NODES.csv')
dirnet = graph_from_data_frame(links,nodes,directed = T)
undirnet = graph_from_data_frame(links,nodes,directed = F)
```

```{r}
plot(dirnet)
```

```{r}
plot(dirnet,vertex.color = 'yellow', vertex.frame.color = 'orange',vertex.shape = 'sphere',vertex.size = 12.5, vertex.label = nodes$media, vertex.label.font = 3 , vertex.label.cex = .75,vertex.label.color = 'black',vertex.label.dist = 1, vertex.label.degree = -pi/2,edge.color = 'red', edge.arrow.size = .25,edge.arrow.width = .5,edge.lty = 'dotted',edge.curved = 1, main = 'Directed Network')
```

```{r}
plot(undirnet)
```

```{r}
plot(undirnet,vertex.color = 'purple', vertex.frame.color = 'blue',vertex.shape = 'square',vertex.size = 10, vertex.label = nodes$media, vertex.label.font = 2 , vertex.label.cex = .75,vertex.label.color = 'black',vertex.label.dist = 1, vertex.label.degree = -pi/2,edge.color = 'blue', edge.lty = 'twodash',edge.curved = 0,main = 'Undirected Network')
```

```{r}
plot(dirnet, vertex.size = nodes$audience.size,edge.width = links$weight,main = 'Directed Network')
plot(undirnet,vertex.size = nodes$audience.size,edge.width = links$weight, main = 'Undirected Network')
```

```{r}
V(dirnet)$size = nodes$audience.size/2
V(undirnet)$size = nodes$audience.size/2
E(dirnet)$width = links$weight/5
E(undirnet)$width = links$weight/5

plot(dirnet, vertex.size = nodes$audience.size/2, vertex.color = 'purple', vertex.label = nodes$media,vertex.label.font = 2, vertex.label.cex = .5, vertex.label.color = 'black',vertex.label.dist = 0,edge.width = links$weight/5,edge.color = 'blue', edge.arrow.width = 0, main = 'Directed Network' )
plot(undirnet, vertex.size = nodes$audience.size/2, vertex.color = 'purple', vertex.label = nodes$media,vertex.label.font = 2, vertex.label.cex = .5, vertex.label.color = 'black',vertex.label.dist = 0,edge.width = links$weight/5,edge.color = 'blue', edge.arrow.width = 0, main = 'Undirected Network' )
```